---
title: "San Francisco International Airport Passenger Traffic"
author: "Jacques Sham & Charles Siu"
date: '`r format(Sys.time(), "%B %d, %Y")`'
output: pdf_document
geometry: margin=0.5in
---

```{r setup, include = F}
knitr::opts_chunk$set(echo = T)
knitr::opts_chunk$set(fig.width = 10)
```

```{r libraries, include = F, warning = F, message = F}
suppressMessages(library(tidyverse))
suppressMessages(library(magrittr))
library(zoo)
library(ggplot2)
library(scales)
library(maps)
library(mapdata)
library(grid)
library(treemapify)
```

```{r init, include = F}
## Set the working dir according to who is working
curr_wd <- strsplit(getwd(), "/")[[1]]
if ("chunheisiu" %in% curr_wd) {
  setwd("/Users/chunheisiu/Dropbox/Documents/USF/2018_Spring/BSDS_100/USF-BSDS100-CaseStudy")
} else if ("jacquessham" %in% curr_wd) {
  setwd("/Users/jacquessham/Documents/GitHub/USF-BSDS100-CaseStudy")
} else {
  wd <- readline(prompt = "You're not Jacques or Charles! Enter the path to the data: ")
  setwd(wd)
}

## Read the data
data <- suppressMessages(read_csv("Air_Traffic_Passenger_Statistics.csv"))
```

## Part 1: Executive Summary
San Francisco is the Pacific gateway to the United States. Commuters travel in and out of the city, and a great percentage is by air, through the San Francisco International Airport (SFO). Maintenance for SFO accounts for a considerable amount in tax expenditure. We would like to investigate whether tax revenue is spent wisely by examining the passenger traffic of SFO. The Government of City and County of San Francisco maintains an open source dataset on the passenger count for every airline that operates at SFO. The dataset contains more than 17,000 entries between 2005 and 2017 on passenger count. Each entry contains the name of the airline, the destination or origin region of the flight, as well as the passenger count. We analyzed the dataset and find out the passenger traffic was growing in a steady rate in the period, increase demand of international passengers with low cost carrier, United Airlines is the largest operator in SFO. We believe passenger traffic determines the effective use of tax revenue in SFO and understand passenger traffic pattern helps decision-making on investment. This analysis allows San Francisco taxpayers to understand how tax is spent on air travel infrastructure, and for airport authorities to propose investments for airport expansion.

## Part 2: Introduction to the Data
### I. Overview
San Francisco is the 13th largest city in the United States and is the Pacific Gateway to the nation. The city's airport SFO, accounts for the majority of air traffic in and out of the SF Bay Area. SFO has been in operation since 1927, with its first international service started in 1946. The airport is the 2nd busiest airport in California, the 7th busiest airport in the United States, and the 23rd busiest airport in the world. SFO has received numerous awards from various air travel rating agencies, such as the 3rd Best Airport in North America in 2012 and 3rd Best Airport Worldwide in 2014 from SkyTrax. There are 3 domestic terminals and 1 international terminal to accommodate the massive flow of passengers travelling to different parts of the United States and the world. 

The data we downloaded from the San Francisco government includes destination, origin, airlines, terminals, and passenger count between July 2005 and December 2017. The data set consists of ``r format(nrow(data), big.mark = ",")`` rows and ``r format(ncol(data), big.mark = ",")`` columns. Here is the link to the dataset: https://data.sfgov.org/Transportation/Air-Traffic-Passenger-Statistics/rkru-6vcg

The report is going to investigate the following facts about SFO:

1) Average monthly passengers traffic between 2006 and 2017
2) Passengers traffic by destination/origin regions
3) Overview on passengers traffic by domestic airlines
4) Passengers traffic travelled by Low Cost Carrier
5) Passengers traffic in airport terminals
6) Passengers traffic on 1 selected domestic carrier

Before we analyze the dataset, we have to clean the data and double check its accuarcy.

```{r cleanup, include = F}
# First, let's clean the data.
# Rename the column names
names(data) <- c("date", "operAirline", "operCode", "airline", "code", "isDomestic",
                 "region", "type", "category", "terminal", "area", "pax")

# Drop operating airline and code columns since they are insignificant
data %<>%
  select(-c(operAirline, operCode))

# Convert isDomestic to boolean
data$isDomestic %<>% recode("Domestic" = T, "International" = F)

# Reformat the dates into Date objects
data$date %<>% 
  as.character() %>% 
  as.yearmon("%Y%m") %>% 
  as.Date()

# Get month and year values
data %<>% 
  mutate(month = date %>% format("%m") %>% factor(labels = month.name),
         year = date %>% format("%Y") %>% as.numeric())

# Remove data from 2005 for easy comparison
data %<>% filter(year != 2005)

data$region %<>% recode("Central America" = "Latin America",
                        "South America" = "Latin America")
data$category %<>% recode("Other" = "Full Service") 

# Change terminal into factor
data %<>%
  mutate(terminal = factor(terminal, levels = c("Terminal 1", "Terminal 2", "Terminal 3", "International", "Other")))
```

We have renamed the column names of the dataset and converted the data structure of some columns for the convenience of the analysis. Here is the explaination of each column:

* **airline** [character]: Airline name
* **code** [character]: Airline Code, code for variable "airline"
* **isDomestic** [logical]: If the flight is `domestic`,T; if `international`: F, flights from/to Canada counts as International
* **region** [character]: Geom region including `US`, `Canada`, `Mexico`, `Latin America`, `Europe`, `Middle East`, `Asia`, and `Australia / Oceania`
* **type** [character]: activity type; `Deplaned` means arrival, `Enplaned` means departure, `Thru / Transit` means flight transit at SFO 
* **category** [factor]: Airline price type; `Low fare` is Low cost carrier, else are others
* **terminal** [factor]: SFO terminal
* **area** [character]: area within SFO terminal
* **pax** [int]: passenger count of given row
* **month** [factor]: The `month` of the airline operates
* **year** [double]: The `year` of the airline operates

We also have to vertify the accuracy of the data, below is how we have cleaned the data.
Below are the steps we took to fix the dataset:

**Step 1: Standardize `United Airlines`**

`United Airlines` merged with `Continent Airlines` in 2013 that some data from `United Airlines` are written in `United Airlines - Pre 07/01/2013` that we have to convert this to `United Airlines` as we can tell those flights were `United Airlines`' operation.
```{r cleanup2_step1, include = F}
# Merge "United Airlines - Pre 07/01/2013" to "United Airlines"
data$airline %<>% recode("United Airlines - Pre 07/01/2013" = "United Airlines")
```

**Step 2: Standardize `Emirates`**

Some of `Emirates` data were written followed with a space. It looks like a typing error but it creates error when analyzing the data, so that I have to standardize all `Emirates` related data.
```{r cleanup2_step2, include = F}
# Remove the strip of "Emirates "
data$airline %<>% recode("Emirates " = "Emirates")
```

**Step 3: Re-identify the category**

There are a lot of `full service airlines` idenify as `low cost carrier`, and vice versa. The international official aviation organization, International Civil Aviation Organization (ICAO),  has an official definition on `low cost carrier` and provides a list of `low cost carrier`. We checked the airlines in the dataset and convert some of the wrongly identified airlines according to he their list. For example, the dataset was wrongly identify `Wow Air` as `full service carrier` but the airline is identify as `low cost carrier` by ICAO.
  
This is the list of low cost carrier: https://www.icao.int/sustainability/Documents/LCC-List.pdf
  
```{r clean2_step3, include = F}
# Some airlines were wrongly identified in category
# Convert the below airlines to full service/Low Fare
full_svc_airline <- c("Air China", "Air India Limited", "Air New Zealand", "Air Pacific Limited dba Fiji Airways",
                      "Emirates", "United Airlines", "Virgin America", "Volaris Airlines", "Delta Air Lines",
                      "US Airways")

lcc_svc_airline <- c("XL Airways France", "WOW Air", "WestJet Airlines")

data %<>%
  mutate(cat_temp = ifelse(airline %in% full_svc_airline, "Full Service", 
                    ifelse(airline %in% lcc_svc_airline, "Low Fare", as.character(category)))) %>% 
  mutate(category = as.factor(cat_temp))
```

## Part 3: Exploratory Analysis

```{r global_vars, echo = F}
## Define 1 Million
million <- 1000000
```

```{r global_theme, echo = F}
## Set a global default on theme and legend on the ggplots
format_title <- theme(plot.title = element_text(hjust = 0.5, size = 16, face = "bold"))
format_legend_title <- theme(legend.title = element_text(face = "bold"))
```

### I. Overview on Passenger traffic in SFO
#### Overall passenger traffic between 2006 and 2017

```{r pax_data, echo = F}
pax_year <- data %>% 
              group_by(year) %>% 
              summarise(sumpax = sum(pax))
```
```{r annual_growth_helper, echo = F}
## Helper function to calculate annual growth rate
growth_rate <- function(x){
  rate <- x / lag(x) - 1
  return (rate)
}
growth_rate_year <- pax_year %>% 
                      mutate(growth = growth_rate(sumpax))
```
```{r section1_data, echo = F}
## Hard fact on passenger growth
max_growth <- growth_rate_year %>% top_n(1,growth)
avg_growth <- paste(round(mean(growth_rate_year$growth, na.rm = T), 4) * 100, "%", sep = "")
double_growth <- ceiling(70 / (mean(growth_rate_year$growth, na.rm = T) * 100))
```

SFO is a busy airport and is a transpacific geteway on the West Coast of the United States, and the passenger traffic is heavy. There are `r format(as.matrix(pax_year[12,2]),big.mark=",")` passengers traveled in SFO in `2017`, compared to `r format(as.matrix(pax_year[1,2]),big.mark=",")` in `2006`. The average annual passenger growth in the period is `r avg_growth`. It will take `r double_growth` years after 2006 for SFO to double the passenger traffic in `r 2006+double_growth`.

#### Average passenger traffic between 2006 and 2017
The airline industry is a season-sensitive industry: Meaning the passenger traffic varies by seasons. Generally, summer and Christmas holidays are two major peak seasons in the airline industry. Therefore, SFO experiences higher passenger traffic in those periods.

Below is the bar chart of monthly average total passenger count of SFO between 2006 and 2017.

```{r average_month_pax}
data %>%
  group_by(isDomestic, month) %>%
  summarise(avg_pax = round(mean(pax), digit = 0)) %>% 
  ggplot(aes(x = factor(month, labels = month.abb), y = avg_pax, fill = isDomestic)) +
  geom_bar(stat = "identity", alpha = 0.8) +
  theme_minimal() +
  scale_y_continuous(labels = comma) + 
  scale_fill_discrete(name = "Destination", label = c("International","Domestic")) +
  labs(x = "Month", y = "Passengers") +
  ggtitle("Monthly Average Passengers Count") +
  geom_text(aes(label = format(avg_pax, big.mark = ",")), size = 2.75,
            position = position_stack(vjust = 0.5), colour = "white") +
  format_title +
  format_legend_title
```

```{r domestic_pax_helper, echo = F}
domestic_pax <- data %>% group_by(isDomestic) %>% summarise(sumpax = sum(pax)) %>% mutate(prct_pax = sumpax/sum(sumpax))
domestic_pax_result <- paste(round(domestic_pax[which(domestic_pax[1]==TRUE),"prct_pax"]*100,2),"%",sep="")
intl_pax_result <- paste(round(domestic_pax[which(domestic_pax[1]==FALSE),"prct_pax"]*100,2),"%",sep="")
```

As seen on the bar chart, `June`, `July`, `August`, `October`, `December` are the months that SFO experiences relatively higher passenger traffic. Beside `October`, the other high-traffic months are in the summer or Christmas holiday seasons. Interestingly, summer is the only busiest period of international traveling in or out from SFO as `June`, `July`, `August` are the only months that SFO handles more than `12000` international travelers. 
About `r domestic_pax_result` of passengers via SFO were domestic travlers. 

### II. Destination
As the bar chart on the monthly average passengers count shows that `r domestic_pax_result` of the SFO travelers are domestic travelers; the remaining `r intl_pax_result` travelers are international travelers.

And the map below shows the passengers count on other regions outside North America between `2006` and `2017`.

```{r map_setup}
# Load the world map and cities
world <- map_data("world")
cities <- world.cities
# Retrieve the information of SF
sf <- world.cities %>%
  filter(name == "San Francisco" & country.etc == "USA")
# Derive the cities and corresponding regions
cities %<>%
  filter(
    (name == "Adelaide" & country.etc == "Australia") | (name == "La Paz" & country.etc == "Bolivia") |
    (name %in% c("Saint Louis", "La Ronge","Riyadh", "Mexico City", "Shenzhen", "Ostrava"))) %>% 
  mutate(region =
           ifelse(name == "Adelaide", "Australia / Oceania", ifelse(name == "Saint Louis", "US",
           ifelse(name == "La Ronge", "Canada", ifelse(name == "La Paz", "Latin America",
           ifelse(name == "Riyadh", "Middle East", ifelse(name == "Mexico City", "Mexico",
           ifelse(name == "Shenzhen", "Asia", ifelse(name == "Ostrava", "Europe", NA))))))))
  ) %>% 
  full_join(y = data %>%
              mutate(region = as.character(region)) %>% 
              group_by(region, type) %>%
              summarize(pax = sum(pax)), by = "region") %>% 
  mutate(
    origin.lat = ifelse(type == "Enplaned", sf$lat, lat),
    origin.long = ifelse(type == "Enplaned", sf$long, long),
    dest.lat = ifelse(type == "Enplaned", lat, sf$lat),
    dest.long = ifelse(type == "Enplaned", long, sf$long)
  ) %>% 
  filter(type != "Thru / Transit")
cities_na <- cities %>% filter(region %in% c("US", "Canada", "Mexico"))
cities_intl <- cities %>% filter(!(region %in% c("US", "Canada", "Mexico")))
```

```{r draw_map}
# Geom objects for drawing static objects
draw_sf_point <- geom_point(x = sf$long, y = sf$lat, color = "red", size = 3)
draw_intl_sf_label <- geom_text(aes(x = sf$long, y = sf$lat, label = "SFO"),
                           hjust = 1, nudge_x = -5, color = "red", size = 3)
draw_color_legend <- scale_color_discrete(name = "Activity Type",
                                          labels = c("Arriving SFO", "Departing SFO"))
draw_size_legend <- scale_size_continuous(trans = "log10", guide = F)

# Functions for drawing curves, points and labels
draw_flight_curve <- function(d) {
  geom_curve(data = d %>% filter(type %in% c("Deplaned", "Enplaned")),
                           aes(x = origin.long, y = origin.lat, xend = dest.long, yend = dest.lat,
                               color = type, size = pax), curvature = 0.5, lineend = "round",
                           alpha = 0.75, arrow = arrow(length = unit(0.025, "npc")))
}
draw_city_points <- function(d) {
  return (geom_point(data = d, aes(x = long, y = lat), color = "black", size = 3))
}

draw_intl_city_labels <- geom_text(data = cities_intl, 
                              aes(x = long, y = lat, label = region),
                              hjust = 0, nudge_x = 3.5, nudge_y = -1,
                              color = "black", size = 3)
draw_intl_enplaned_labels <- geom_label(data = cities_intl %>% filter(type == ("Enplaned")),
      aes(x = long, y = lat, label = format(pax, big.mark = ","), color = type),
      hjust = 0, nudge_x = 3.5, nudge_y = -7, size = 3, show.legend = F)
draw_intl_deplaned_labels <-geom_label(data = cities_intl %>% filter(type == ("Deplaned")),
      aes(x = long, y = lat - 4, label = format(pax, big.mark = ","), color = type),
      hjust = 0, nudge_x = 3.5, nudge_y = -9.5, size = 3, show.legend = F)

format_theme <- theme(
  axis.text = element_blank(),
  axis.line = element_blank(),
  axis.ticks = element_blank(),
  panel.border = element_blank(),
  panel.grid = element_blank(),
  axis.title = element_blank(),
  legend.position = "bottom",
  legend.background = element_rect(fill = "gray90", size = 0),
  legend.title = element_text(face = "bold")
  )
```
```{r draw_intl_map, fig.height = 8}
world %>% 
  filter(region != "Antarctica") %>% 
  ggplot() +
  geom_polygon(aes(x = long, y = lat, group = group), fill = "darkgray") +
  draw_flight_curve(cities_intl) +
  draw_city_points(cities_intl) +
  draw_intl_city_labels +
  draw_sf_point +
  draw_intl_sf_label +
  draw_intl_enplaned_labels +
  draw_intl_deplaned_labels +
  coord_fixed(1.3) +
  theme_minimal() +
  draw_color_legend +
  draw_size_legend +
  scale_x_continuous(limits = c(-170, 200)) +
  scale_y_continuous(limits = c(-60, 90)) +
  ggtitle("Passengers Count by International Destinations") +
  format_theme +
  format_title
```
```{r region_pax}
region_pax <- data %>% group_by(region) %>% summarise(sumpax = sum(pax)) %>% arrange(desc(sumpax))
```

The map shows that `r region_pax[2,1]` is the continent of the most passengers flying from or to after the United States, with `r format(as.matrix(region_pax[2,2]), big.mark=",")` passengers, followed by `Europe`. Surprisingly there are less passengers coming from or going to `Canada` and `Mexico` than they coming from or going to `Asia` or `Europe`, `r format(as.matrix(region_pax[4,2]), big.mark=",")` and `r format(as.matrix(region_pax[5,2]), big.mark=",")` passengers, respectively, and the passenger traffic coming from or going to `Latin America` is very little, with `r format(as.matrix(region_pax[8,2]), big.mark=",")`. 

### III. Domestic Carriers Overview
There are a lot of operators in SFO. Below is the stacked line chart of the domestic passengers traffic by airlines.

```{r top5_dom_list}
# Filter the top 5 airlines by domestic passenger count
top5_dom_list <- data %>%
  filter(isDomestic) %>%
  group_by(airline) %>%
  summarise(total_pax = sum(pax)) %>%
  top_n(5, total_pax) %>%
  arrange(total_pax) %>% 
  select(-total_pax)
# Combine and compute the other airlines
other_dom_airline <- data %>% 
  filter(!(airline %in% top5_dom_list$airline)) %>% 
  group_by(year) %>% 
  summarise(sum = sum(pax)) %>% 
  mutate(airline = "Other Airlines") %>% 
  select(airline, year, sum)
```

```{r top5_dom_area}
data %>%
  group_by(airline, year) %>% 
  summarize(sum = sum(pax)) %>% 
  right_join(top5_dom_list, by = "airline") %>%
  ungroup() %>% 
  rbind(other_dom_airline) %>%
  rbind(data %>%
          group_by(airline) %>%
          right_join(top5_dom_list, by = "airline") %>%
          summarize(year = min(year) - 1) %>%
          filter(year == min(data$year)) %>%
          mutate(sum = 0)) %>% 
  mutate(airline = factor(airline, 
      levels = rbind(top5_dom_list, "Other Airlines")$airline)) %>% 
  ggplot() +
  geom_area(aes(x = year, y = sum / million, fill = airline), alpha = 0.75) +
  scale_x_continuous(name = "Year",
          breaks = seq(min(data$year), max(data$year), by = 1)) +
  scale_y_continuous(name = "Passengers (Millions)") +
  scale_fill_brewer(name = "Airline", palette = "Set2") +
  theme_minimal() +
  ggtitle("Domestic Passengers Count by Airline") +
  format_title +
  format_legend_title
```
```{r dom_travel_info, echo = F}
airline_dom_pax <- data %>%
                      filter(isDomestic) %>%
                      group_by(airline) %>%
                      summarise(total_pax = sum(pax))
top5_dom_pax <- airline_dom_pax %>% 
                  top_n(5, total_pax) %>% 
                  summarise(pax = sum(total_pax))
dom_pax <- airline_dom_pax %>% summarise(pax = sum(total_pax))
```
```{r travel_fact, include = F}
top5_dom_pax_num <- top5_dom_pax[1,1]
all_dom_pax <- dom_pax[1,1]
non_top5_dom_pax <- dom_pax[1,1] - top5_dom_pax[1,1]
top5_dom_portion<- paste(round(top5_dom_pax_num/all_dom_pax,4)*100,"%",sep="")
non_top5_dom_portion <- paste(round(non_top5_dom_pax/all_dom_pax,4)*100,"%",sep="")
```
The chart shows that a large portion of SFO travelers are traveling by `r top5_dom_list[5,1]`, while SFO is one of the hub of `r top5_dom_list[5,1]` in the West Coast. `r top5_dom_list[1,1]`, `r top5_dom_list[2,1]`(Recently Merged with `Alaska Airlines` in 2017), `r top5_dom_list[3,1]`, `r top5_dom_list[4,1]` are the other top 5 major carrier in SFO in passenger count. A large portion of travelers in SFO travel with those airlines, `r top5_dom_portion` of all domestic travelers traveled with those airlines between `2006` and `2017`. 

###III. Low Cost Carrier vs Full Service Carrier
Due to the advance of avaition technology, the cost of air travel significantly decrease in the last 40 years. Low cost carriers provide low cost traveling by reducing significant ticket price to attract passengers; to make up the loss of revenue, `low cost carriers` cut cost on meals, seating comfort, onboard entertainment, and seating priority. Alternatively, airlines that provide decent meals, onboard entertainment, and offer comfortable seats and seating priority are called `full Service Carrier`. For travelers travel without too much concern on those feature, traveling with low cost carrier is a good choice for them. `Southwest Airlines` is one of the successful story and is a low cost carrier giant in the United Staets. In 2016, Iceland-based low cost carrier, `Wow Air`, began scheduled service from Reykjavik, Iceland to SFO that makes `Wow Air` the first low cost carrier operates regular scheduled service to SFO outside of the North America.

```{r lcc_info, include = F}
lcc_pax <- data %>% filter(year==2017) %>% 
  group_by(category) %>% summarise(sumpax = sum(pax)) %>% 
  mutate(prct_pax = sumpax/sum(sumpax))

lcc_pax_prct <- paste(round(lcc_pax[2,"prct_pax"],4)*100,"%",sep="")
lcc_pax_count <- lcc_pax[2,"sumpax"]

southwest_count <- data %>% filter(airline=="Southwest Airlines") %>% group_by(year) %>% summarise(yearpax = sum(pax))

southwest_pax_growth <- southwest_count %>% filter(year >= 2010) %>% mutate(growth = growth_rate(yearpax))
southwest_pax_y_growth <- southwest_pax_growth  %>% summarise(mean_growth = mean(growth, na.rm=T))

wowair_pax_growth <- data %>% filter(airline=="WOW Air") %>% 
  group_by(year) %>% summarise(yearpax = sum(pax)) %>% 
  mutate(growth = growth_rate(yearpax)) %>% summarise(mean_growth = mean(growth, na.rm=T))
```

In 2017, about `r format(as.matrix(lcc_pax_count[1,1]),big.mark=",")` passengers traveled via SFO by low cost carriers, which makes up about `r lcc_pax_prct` of the total passengers traffic.

```{r lcc_basic, include = F}
intl_lcc_plot <- data %>% 
                  filter(!isDomestic & 
                           category=="Low Fare" &
                           airline != "ATA Airlines" &
                           airline != "Servisair" &
                           airline != "Sun Country Airlines") %>% 
                  group_by(year,airline) %>% 
                  summarise(sumpax = sum(pax)) %>% 
                  arrange(desc(sumpax))
```

Prior to 2007, not a lot of passengers traveled by low cost carriers via SFO. The passengers traffic skyrocketed between 2007 and 2009 due to low cost carrier giant `Southwest Airlines` had began service in SFO. `Southwest Airlines` served `r format(as.matrix(southwest_count[1,2]),big.mark=",")`, `r format(as.matrix(southwest_count[2,2]),big.mark=",")`, `r format(as.matrix(southwest_count[3,2]),big.mark=",")` passengers in `r southwest_count[1,1]`, `r southwest_count[2,1]`, and `r southwest_count[3,1]`, respectively. After 2009, the annual passenger growth `r paste(round(southwest_pax_y_growth,4)*100,"%",sep="")`. In `r southwest_count[11,1]`, `r format(as.matrix(southwest_count[11,2]),big.mark=",")` passengers traveled with `Southwest Airlines`.

There was no international low cost carrier provide regular scheduled service to SFO prior to 2016, `WestJet Airlines` and `XL Airways France` the only international low cost carrier operate seasonal routes in SFO. In 2016, `Wow Air` annouced began scheduled service from Reykjavik, Iceland. Below is the bar chart on the passengers count on international low cost carriers. 

```{r intl_lcc_barchart}
intl_lcc_plot %>% ggplot(aes(x = year, y = sumpax, fill = airline)) +
  scale_fill_brewer(palette = "Spectral") +
  geom_bar(stat = "identity") + 
  scale_x_continuous(name = "Year",
        breaks = seq(min(data$year), max(data$year), by = 1)) +
  scale_y_continuous(name = "Passengers", breaks = seq(0, 500000, by = 50000), labels = comma)+
  ggtitle("The Annual Passengers Count on International Low Cost Carrier")+
  theme_minimal() +
  format_title +
  format_legend_title
```

The passengers traffic almost double when `Wow Air` between 2016 and 2017, the first year began service to SFO, it was `r paste(round(wowair_pax_growth,4)*100,"%",sep="")`. However, `Wow Air`'s service to SFO is new, we expect the growth rate would smooth out to single digit by 2019 if we observe the pattern from `Southwest Airlines`' first 3 years of service in SFO. 

Although many travelers do not travel with low cost carriers, there are increasingly more passengers travel with `Southwest Airlines` and `Wow Air`. Therefore, we expect the portion of passengers traveled with low cost carrier increase. 

### IV. Terminal Traffic
There are 4 terminals in SFO along with 115 gates: Terminal 1 (T1), Terminal 2 (T2), Terminal 3(T3), and International Terminal(IT). T1, T2, and T3 are designed to handle domestic and precleared flights from Canada, and IT are to handle international flgiths. 

Below is the tree map on domestic passenger traffic by terminal in 2017.

```{r terminal_domestic}
data %>%
  filter(isDomestic, !is.na(code) & year == 2017) %>%
  group_by(terminal, airline, code) %>%
  summarise(all_pax = sum(pax)) %>% 
  ggplot(aes(area = all_pax, fill = terminal, label = code, group = airline)) +
  geom_treemap() +
  geom_treemap_text(colour = "white", place = "centre") +
  scale_fill_brewer(name = "Terminal", palette = "Set2") +
  ggtitle("Domestic Passengers Count by Airline and Terminal") +
  format_title +
  format_legend_title
```
```{r domestic_pax}
pax_terminal <- data %>% filter(isDomestic & year == 2017 & terminal != "other") %>% 
                  group_by(terminal) %>% 
                  summarise(sumpax = sum(pax)) %>% 
                  mutate(prct_pax = sumpax/sum(sumpax)) %>% 
                  arrange(desc(prct_pax))
```

As we can see from the tree map: `United Airlines` is assigned to dock at T3, although some `United Airlines` flights dock at IT. `Delta Airlines`, `Southwest Airlines`, `Frontier Airlines` and some `American Airlines` flights were docking at T1. T2 serves the flights of `Alaska Airlines`, `Virgin Airlines`, and the remaining `American Airlines` flights. Due to the constraint capcity of T1, T2, T3, the remaining domestic carriers with less flight frequency from/to SFO, including `Hawaiian Airlines`, `Jetblue Airways`, `Sun Country Airlines`, and some `Alaska Airlines` flights, are docking in IT even those are domestic flights.

About `r paste(round(pax_terminal[1,3],4)*100,"%",sep="")` passengers traveled via `r pax_terminal[1,1]` in 2017 that makes `r pax_terminal[1,1]` the busiest terminal in SFO, followed by `r pax_terminal[2,1]` with `r paste(round(pax_terminal[2,3],4)*100,"%",sep="")` of the domestic passengers. The remaining domestic passenger traveled in `r pax_terminal[3,1]` and `r pax_terminal[4,1]`, those make up of `r paste(round(pax_terminal[3,3],4)*100,"%",sep="")` and `r paste(round(pax_terminal[4,3],4)*100,"%",sep="")` of the domestic passenger traffic, respectively.

Almost half of the domestic passengers in 2017 traveled with `United Airlines` which makes up the largest share of the domestic flight market, followed by the major domestic airlines includes `American Airlines`, `Alaska Airlines`, `Delta Airlines`, `Virgin America`, and the low cost carrier giant `Southwest Airlines`

### VI. Airlines

`United Airlines` is largest carrier in SFO in terms of passengers traffic in 2017, both in domestic or international flights since `United Airlines` set SFO as one of its hub; `United Airlines` has assigned a lot of flights fly in and out from SFO. In `United Airlines`' prespective, SFO is the 5th largest hub measured by number of flights, and the primary hub on the West Coast.

`United Airlines`' footstep in SFO can be traced back in 1937, `United Airlines` operated scheduled service to Los Angeles and New York in January 1937 after it was formed in 1934. And `United Airlines` has one of the largest single aircraft maintence bases in SFO. 

The below heated map show the passenger traffic between 2007 and 2017 for `United Airlines`:

```{r united}
data %>%
  filter(code == "UA") %>%
  group_by(month, year) %>%
  summarise(Passengers = sum(pax)) %>%
  ggplot(aes(x = factor(month, labels = month.abb), y = year)) +
  geom_tile(aes(fill = Passengers)) + 
  scale_x_discrete(name = "Month") +
  scale_y_continuous(expand = c(0, 0),
    name = "Year", breaks = seq(min(data$year), max(data$year), by = 1)) +
  scale_fill_gradientn(colours = rev(heat.colors(10)), labels = comma) +
  theme_minimal() +
  ggtitle("United Airlines Passengers Count") +
  format_title +
  format_legend_title
```

The above heated map shows the passenger traffic is relatively low by 2012; the passenger traffic jumped after 2012 and continue to grow. Also, we can also confirm that there are more passengers travel by `United Airlines` in summer period as `June`, `July`, and `August` of every year tend to have a darker colored pattern compare to the months of the same years, that means `United Airlines`' operation is very sensitive to seasonal effect.

## Part 4: Conclusions and Future Analysis

In the report, we found out that the passenger traffic growth was continuous between `2006` and `2017` and we expect that this growth will continue. In this report, we found out `r domestic_pax_result` of the passengers are domestic travelers and most of the domestic travelers travel with `r top5_dom_list[5,1]`. There are increasingly more travelers who travel out of the United States by low cost carriers. The largest operator in SFO, `United Airlines` is very sensitive to seasonal effect. After we have received the result from the analysis, we can see that the tax revenue was spent reasonable on air travel infrastructure in San Francisco because the passenger traffic was growing between the time period, realize the need for long-term investment in SFO to increase the capcity for passenger travel to serve more travelers in the future. At the same time, SFO adminstrative staffs can accommodate the short-term high demand on the airport service in the summer period, `June`, `July`, `August` after learning from the analysis.

However, the dataset we have for this report focus mainly on the passenger counts of region the flights fly to or from, there was no information on destination/origin cities and country, aircraft models. In this report; we are not able to trace any of destination/origin cities and country of the airlines. In the future, we would like to explore more on destination/origin cities and country and aircraft models each flight. Additionally, we are surprised on the passenger traffic of `United Airlines` continue to grow although the `United Airlines`' reputation worsened in recent years. We would like to observe the passenger traffic change on `United Airlines` in the next few years if the reputation effect occurs.